---
title: 'Exercise 5: Functions'
author: "Marcy Shieh"
date: "9/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```

# Prerequisites

1. Open you `ps811-exercises` folder.

2. Go to File > New File > R Script.

3. Save file as `exercise-6-functions.R`.

4. Load R packages installed in Lectures 5 and 6.

```{r iv, eval = FALSE}
library(tidyverse)
library(here)
```

5. Download the [national.csv file from Kaggle](https://www.kaggle.com/umichigan/world-religions).

6. Load the CSV file into your R environment.

```{r vi}
national <- read.csv(here("data/national.csv"))
```

7. View the CSV.

# Loops

1. Write a loop to find how many variables there are per observations.

```{r q1}
rows_loop <- numeric()

for (i in 1:ncol(national)) {
  rows_loop[i] <- length(national[i, ])
}

rows_loop
```

2. Find the mean number of Protestant Christians in each country (i.e., the state column) using `tapply()` in both base R and Tidyverse.

```{r q2, eval = FALSE}
# base R
tapply(
  X = national$christianity_protestant, 
  INDEX = list(national$state), 
  FUN = mean, 
  na.rm = TRUE  
)

# tidyverse
national %>%
  group_by(state) %>%
  summarize(
    mean_christ= mean(christianity_protestant, na.rm = TRUE)
  ) 

```

3. Check the column type for each variable that is just characters.

    Hint: You can use `sapply(dataset-name-here, class)`.
    
```{r q3, eval = FALSE}
sapply(national, is.character)
```

4. Log the `buddhism` variables.

```{r q4, eval = FALSE}
national %>%
  mutate_at(
    .vars = vars(starts_with("buddhism")),
    .funs = function(x) log(x)
  ) %>% 
  select(starts_with("buddhism"))

# answers referring to only the "buddhism_all" variable are also acceptable
# I realize that this question is unclear (sorry!)
```

5. Write a function that lists all the unique years with more than 300,000 Christians in total.

```{r q5}
unique_chr300k_yr <- function(x) {
    if(national$christianity_all > 300000) {
return(unique(national$year))
    }
}

unique_chr300k_yr()
```

6. Group by the code variable.

    Hint: Use nesting.
    
```{r q6}
nested_code <- national %>%
  group_by(code) %>%
  nest()
```

7. Create a model column. You need to include a linear regression model that examines whether the proportion of Jews (judaism_percent) affect whether a country has dual religions.

```{r q7}
nested_models <- nested_code %>%
  mutate(
    model = map(
      .x = data, 
      .f = ~ lm(dual_religion ~ judaism_percent, data = .x)
    )
  ) %>%
  print()
```

8. Extract the coefficients into a new column

```{r q8}
nested_coefs <- nested_models %>%
  mutate(coefs = map(model, coefficients)) %>%
  print()
```

9. Look at the coefficients.

```{r q9, eval = FALSE}
nested_coefs$coefs
```

10. Pull out the model column.

```{r q10, eval = FALSE}
nested_coefs %>% pull(model)
```

11. Unnest the coefficients column.

```{r q11}
coefs <- nested_coefs %>%
  unnest(coefs) %>%
  print()
```