---
title: 'Exercise 10: Data Manipulation with Replication Exampl (Answer Key)'
author: "Marcy Shieh"
date: "12/5/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages
library("magrittr")
library("tidyverse")
library("haven")
library("here")
library("stargazer")
library("modelsummary")
library("sjlabelled")

here("data")
```

# Extend the dataset

Extend the exercise we did in class up to the Senate confirmation of Associate Justice Sonia Sotomayor, following the instructions in the `lecture-10.R` file.

Links to the datasets required to complete this exercise:

- Harvard Dataverse entry for the Epstein et al. (2006) article (includes PDF of article for reference): https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/W5CV8E&version=1.0

- Poole & Rosenthal scores from 2016: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/W5CV8E&version=1.0

- 111th Congress roll call votes and dictionary: https://legacy.voteview.com/senate111.htm

- Segal-Cover scores: https://en.wikipedia.org/wiki/Segal%E2%80%93Cover_score

```{r alito, include = FALSE}
bork <- read_dta(here::here("data", "Bork.dta")) # download from https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/W5CV8E&version=1.0

# table 1: descriptive stats - mean, deviation, min, max
# senate vote (vote), lack of qualifications (lackqual), ideological distance (eucldist), strong president (strngprs), same party (sameprty)
# https://vincentarelbundock.github.io/modelsummary/articles/datasummary.html
datasummary((`Senate Vote` = vote) +
              (`Lack of Qualifications` = lackqual) +
              (`Ideological Distance` = eucldist) +
              (`Strong President` = strngprs) +
              (`Same Party` = sameprty) ~
              Mean + SD + Min + Max,
            data = bork,
            output = 'markdown')
# see other display options: https://vincentarelbundock.github.io/modelsummary/articles/datasummary.html#save-and-display-1

# table 2: senate voting on supreme court nominees
# time frame: from black to bork
table2_ccs_nominees <- filter(bork, congress > 82 & congress < 101)
table2_ccs_nominees_model <- glm(vote ~ lackqual + eucldist + strngprs + sameprty,
                                 family = binomial(link = "probit"), 
                                 data = table2_ccs_nominees)
summary(table2_ccs_nominees_model)

# time frame error in the paper: it claims this is the time frame from black to roberts but it seems to be the time frame between black to alito (did not realize they also included alito in their analysis lol)
table2_add_nominees <- glm(vote ~ lackqual + eucldist + strngprs + sameprty,
                           family = binomial(link = "probit"), 
                           data = bork)
summary(table2_add_nominees)

stargazer(table2_ccs_nominees_model, table2_add_nominees,
          covariate.labels = c("Lack of Qualifications", "Ideological Distance",
                               "Strong President",
                               "Same Party"),
          column.labels = c("CCS Nominees", "Additional Nominees"),
          dep.var.caption = "Senate Vote",
          dep.var.labels.include = FALSE,
          type = "text")
# loglik error: notice that the log likelihood for the Additional Nominees are off in the paper

# add alito to the dataset

# so we need the common space score for each senator (dwnom1)
poole <- read_dta(here::here("data", "Weekly_DW-NOMINATE_31.DTA")) # download from https://legacy.voteview.com/Weekly_Constant_Space_DW-NOMINATE_Scores.htm

# there is leading whitespace in poole$name
# let's fix this so the names in both datasets merge properly
poole$name <- trimws(poole$name)

# rename congress
poole <- rename(poole, congress = cong)

# say you want to look at additional nominees
# extension: from black to alito

# you need to extend the dataset
# build alito votes (alito was confirmed on 1/31/2006)
# https://legacy.voteview.com/senate109.htm

# download all the roll call votes of the 109 congress
congress109 <- read_dta(here::here("data", "sen109kh.dta"))

# identify which column contains the alito votes by downloading the dictionary
# it seems like V368 is the alito vote column
congress109_alito <- congress109 %>%
  select(cong:name, V368) %>%
  filter(name != "BUSH") # filter out the president

# look at the breakdown
table(congress109_alito$V368)

# 1 = YES, 6 = NAY, 0 = not a member
# you want to change 6 to 0
# you want to change 0 to NA (CORZINE retired)
congress109_alito <- congress109_alito %>%
  mutate(
    vote = ifelse(V368 == 6,
                  0,
                  ifelse(V368 == 0,
                         NA,
                         V368)
    ))

# check the breakdown again
table(congress109_alito$vote)

# you want to build the alito file to reflect the bork file
# so first, check the variable names

# bork
# "congress"(+)     "statenm"(+)      "name"(+)         "cs1"(+)          "nominee"(-) 
# "vote" (+)         "presprty"(-)     "sameprty"(-)     "strngprs"(-)     "nomid"(-)  
# "regime_dummy"(-) "lackqual"(-)     "csnom"(-)        "eucldist"(-)     "interaction"(-)

names(congress109_alito)
names(bork)

congress109_alito <- rename(congress109_alito, congress = cong)

alito <- congress109_alito %>%
  select(congress, id, state, name, vote)

# select relevant variables in the poole & rosenthal common space scores
poole_select <- poole %>%
  select(congress, idno, statenm, name, dwnom1, party)

# merge the senators' alito votes with the senators' common space scores
alito_poole <- left_join(alito, poole_select, by=c("id" = "idno", "congress", "name"))
names(alito_poole)
alito_poole <- rename(alito_poole, cs1 = dwnom1)
# now you want to build a dataframe with the following information:
# nominee, presprty, sameprty, strngprs, nomid, regime_dummy, lackqual, csnom, eucldist, interaction
alito_poole$nominee <- rep("ALITO") # name of nominee
alito_poole$presprty <- rep(1) # president held party majority in Senate

alito_poole$sameprty <- rep(NA)
# whether senator is the same party as the president
alito_poole %>%
  mutate(sameprty = ifelse(party == 200,
                           1,
                           0))

alito_poole$nomid <- rep(0.100) # ideology score of the judicial nominee: https://en.wikipedia.org/wiki/Segal%E2%80%93Cover_score
alito_poole$regime_dummy <- rep(1) # whether the judicial nominee was pre- or post-bork (1 is post-bork, 0 is pre-bork)
alito_poole$lackqual <- rep(1-0.810) # you take how qualified someone is based on newspaper editorials and subtract it by 1 to get their "lack of qualifications" score #https://en.wikipedia.org/wiki/Segal%E2%80%93Cover_score

# you will need to calculate the following (later)
alito_poole$csnom <- NA # this is going to be the Segal-Cover score for each nominee "transformed" as the Common Space/Poole & Rosenthal/DW-Nominate (it goes by many names) score

alito_poole$eucldist <- NA # this is going to be the Euclidean distance between the senator and the nominee

alito_poole$interaction <-NA # this is the interaction between ideological distance and qualifications

# bind the dataset to the bork dataset
bork_alito_poole <- full_join(bork, alito_poole)
names(bork_alito_poole)
# the last two columns are just for binding datasets before
# you can just keep them there if they don't interfere with your analysis going forward

# from reading the article, you know that you need the common space scores for each president as well
# the president is coded with a congressional district (cd) number of 0, and a state number (state) of 99
poole_filter <- poole %>%
  filter(cd == 0 & state == 99) %>%
  select(congress, name, dwnom1)

# rename name
poole_filter <- rename(poole_filter, pres = name)

# rename dwnom1
poole_filter <- rename(poole_filter, cspres = dwnom1)

# join together the cspres scores
bork_alito_poole_full <- full_join(bork_alito_poole, poole_filter)

alito_poole_coeffs <- lm(cspres ~ nomid, data = bork_alito_poole_full)
alito_poole_coeffs$coefficients
# result:
# (Intercept)       nomid 
# 0.5816398  -0.9868674 

# scale the common space scores for alito
bork_alito_poole_full$csnom <- 0.5816398 - 0.9868674*bork_alito_poole_full$nomid

# do the calculations for ideological distance
bork_alito_poole_full$eucldist <- (bork_alito_poole_full$csnom - bork_alito_poole_full$cs1)^2

# do the calculations for the interaction between ideological distance and qualifications
bork_alito_poole_full$interaction <- bork_alito_poole_full$eucldist*alito_poole$lackqual

# now you have the full dataset so you can run the analysis again from black to alito (and not just black to roberts)
```


```{r sotomayor}
# download 111th congress roll call votes
congress111 <- read_dta(here::here("data", "sen111kh.dta"))

# V263 = Sotomayor senate vote
congress111_sotomayor <- congress111 %>%
  select(cong:name, V262) %>% # it is on row 263 meaning it is V262 (consider the header)
  filter(name != "OBAMA") # exclude Obama (president)

# 1 = YES, 6 = NAY, 0 = not a member
# you want to change 6 to 0
# you want to change 0 to NA 
congress111_sotomayor <- congress111_sotomayor %>%
  mutate(
    vote = ifelse(V262 == 6,
                  0,
                  ifelse(V262 == 0,
                         NA,
                         V262)))

# you can confirm that this is the correct vote column by looking up "sotomayor confirmation" and see that it matches the table below...
table(congress111_sotomayor$vote)

# rename "cong" variable to "congress"
congress111_sotomayor <- rename(congress111_sotomayor, congress = cong)

# select the relevant IDs
sotomayor <- congress111_sotomayor %>%
  select(congress, id, state, name, vote)

# select relevant variables to the poole & rosenthal CS scores
poole_select <- poole %>%
  select(congress, idno, statenm, name, dwnom1, party)

# merge the senate votes with their CS scores
sotomayor_poole <- left_join(sotomayor, poole_select, by=c("id" = "idno", "congress", "name"))

# rename "dwnom1" to "cs1"
sotomayor_poole <- rename(sotomayor_poole, cs1 = dwnom1)

# now you want to build a dataframe with the following information:
# nominee, presprty, sameprty, strngprs, nomid, regime_dummy, lackqual, csnom, eucldist, interaction
sotomayor_poole$nominee <- rep("SOTOMAYOR") # name of nominee
sotomayor_poole$presprty <- rep(1) # president held party majority in Senate

# whether senator is the same party as the president
sotomayor_poole %>%
  mutate(sameprty = ifelse(party == 100,
                           1,
                           0))

sotomayor_poole$nomid <- rep(0.780) # ideology score of the judicial nominee: https://en.wikipedia.org/wiki/Segal%E2%80%93Cover_score

sotomayor_poole$regime_dummy <- rep(1) # whether the judicial nominee was pre- or post-bork (1 is post-bork, 0 is pre-bork)

sotomayor_poole$lackqual <- rep(1-0.810) # you take how qualified someone is based on newspaper editorials and subtract it by 1 to get their "lack of qualifications" score #https://en.wikipedia.org/wiki/Segal%E2%80%93Cover_score

# you will need to calculate the following (later)

sotomayor_poole$csnom <- NA # this is going to be the Segal-Cover score for each nominee "transformed" as the Common Space/Poole & Rosenthal/DW-Nominate (it goes by many names) score

sotomayor_poole$eucldist <- NA # this is going to be the Euclidean distance between the senator and the nominee

sotomayor_poole$interaction <-NA # this is the interaction between ideological distance and qualifications

# bind the dataset to the bork dataset
all_data <- full_join(bork_alito_poole, sotomayor_poole)
# the last two columns are just for binding datasets before
# you can just keep them there if they don't interfere with your analysis going forward

# from reading the article, you know that you need the common space scores for each president as well
# the president is coded with a congressional district (cd) number of 0, and a state number (state) of 99
poole_filter <- poole %>%
  filter(cd == 0 & state == 99) %>%
  select(congress, name, dwnom1)

# rename name
poole_filter <- rename(poole_filter, pres = name)

# rename dwnom1
poole_filter <- rename(poole_filter, cspres = dwnom1)

# join together the cspres scores
all_data_full <- full_join(all_data, poole_filter)

sotomayor_poole_coeffs <- lm(cspres ~ nomid, data = all_data_full)
sotomayor_poole_coeffs$coefficients
# result:
# (Intercept)       nomid 
# 0.5832055  -0.9999904 

# scale the common space scores for sotomayor
all_data_full$csnom <- 0.5832055 - 0.9999904*all_data_full$nomid

# do the calculations for ideological distance
all_data_full$eucldist <- (all_data_full$csnom - all_data_full$cs1)^2

# do the calculations for the interaction between ideological distance and qualifications
all_data_full$interaction <- all_data_full$eucldist*sotomayor_poole$lackqual

# now you have the full dataset so you can run the analysis again from black to sotomayor (and not just black to roberts)
```

```{r data summary}
datasummary((`Senate Vote` = vote) +
              (`Lack of Qualifications` = lackqual) +
              (`Ideological Distance` = eucldist) +
              (`Strong President` = strngprs) +
              (`Same Party` = sameprty) ~
              Mean + SD + Min + Max,
            data = all_data_full,
            output = 'markdown')
```

```{r include = FALSE}
# table 2: senate voting on supreme court nominees
# time frame: from black to bork
table2_ccs_nominees <- filter(bork, congress > 82 & congress < 101)
table2_ccs_nominees_model <- glm(vote ~ lackqual + eucldist + strngprs + sameprty,
                                 family = binomial(link = "probit"), 
                                 data = table2_ccs_nominees)
summary(table2_ccs_nominees_model)

# time frame error in the paper: it claims this is the time frame from black to roberts but it seems to be the time frame between black to alito (did not realize they also included alito in their analysis lol)
table2_add_nominees <- glm(vote ~ lackqual + eucldist + strngprs + sameprty,
                           family = binomial(link = "probit"), 
                           data = bork)
summary(table2_add_nominees)

table2_newadditions <- glm(vote ~ lackqual + eucldist + strngprs + sameprty,
                           family = binomial(link = "probit"), 
                           data = all_data_full)
```

```{r, results='asis'}
stargazer(table2_ccs_nominees_model, table2_add_nominees, table2_newadditions,
          covariate.labels = c("Lack of Qualifications",
                               "Ideological Distance",
                               "Strong President",
                               "Same Party"),
          column.labels = c("CCS Nominees",
                            "Additional Nominees",
                            "Add Alito and Sotomayor"),
          dep.var.caption = "Senate Vote",
          header = FALSE,
          dep.var.labels.include = FALSE,
          type = "latex")
```

# Replication

Find replication materials for a paper and replicate a table within the paper.

- There are many political science papers with data stored in the Stata (`.dta`) file  format accompanied by a Stata file (`.do`) file. In those cases, you will need to read the Stata file into R using the `haven` package and recreate your table of choice using R code. In other words, please do not turn in Stata code!
    
- If you do find replication materials in R and would like to run the entire R script, please feel free to do so! Reviewing other people's replication materials is the best way to learn how scholars assemble their results.

Here are some great resources for finding articles with reproducible data.

- Political Science Replication (https://politicalsciencereplication.wordpress.com/)

- Harvard Dataverse (https://dataverse.harvard.edu/dataverse/harvard?q=&fq0=subject_ss%3A%22Social%20Sciences%22&types=dataverses%3Adatasets&sort=dateSort&order=desc)

- Recent articles from *Political Science Research and Methods*, *American Journal of Political Science*, *American Political Science Review*, and *Political Analysis*

    - You can access articles in the top journals by signing into the UW-Madison Library website (https://library.wisc.edu) with your UW netID

# Submit

Email me (mshieh2@wisc.edu) the link to your `ps811-exercises` repository when you are done.