---
title: "Lecture 8. Graphics"
author: Marcy Shieh
date: October 29, 2020
output: beamer_presentation
---

# Goals

- Learn how to create quick graphs
- Learn how to create pretty graphs

# Why graphs?

- Graphs help you and your readers visualize your data
- Use graphs for your presentations
    - I tend to break this rule a lot because I'm terrible and like regression tables
    - But in general, people like it when you use graphs
- Use graphs for your papers
    - Put the table in the appendix!
    
# base R graphs

- These are kinda ugly (no offense)
- But they are useful when you just need to...see something...as a graph
- Think of base R graphs as a sketchpad. You can use it to draft something but do not use it as your final version (IMO).

# base R graph example

```{r base-R, echo = FALSE}
mtcars$cyl <- as.factor(mtcars$cyl)
plot(mpg ~ hp , data=mtcars, 
    col=cyl, pch=c(4,6,8)[mtcars$cyl], cex=1.2)
legend("topright", legend=levels(mtcars$cyl),
                   pch = c(4,6,8),
                   col=levels(mtcars$cyl))
```

# `ggplot2` graphs

- `ggplot2` graphs are prettier
- Package has more themes and allows you to customize your graphs in a billion different ways
- And unlike base R graphs, `ggplot2` graphs look like they were made in the 21st century

# `ggplot2` graph example

```{r ggplot2, echo = FALSE}
library(ggplot2)
ggplot(mtcars, aes(x=hp, y=mpg, color=cyl, shape=cyl)) +
    geom_point(size=3)
```
