\documentclass[11pt,]{article}
\usepackage[margin=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}
\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center
\newcommand{\blankline}{\quad\pagebreak[2]}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}} 
\usepackage{longtable,booktabs}

\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}

\titleformat*{\subsubsection}{\normalsize\itshape}

\usepackage{titling}
\setlength{\droptitle}{-.25cm}

%\setlength{\parindent}{0pt}
%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%\setlength{\emergencystretch}{3em}  % prevent overfull lines 

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{lastpage}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0.0pt} 
\lhead{}
\chead{}
\rhead{\footnotesize DRAFT - PS 811: Introduction to Statistical Computing -- Fall 2020}
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\fancypagestyle{firststyle}
{
\renewcommand{\headrulewidth}{0pt}%
   \fancyhf{}
   \fancyfoot[C]{\small \thepage/\pageref*{LastPage}}
}

%\def\labelitemi{--}
%\usepackage{enumitem}
%\setitemize[0]{leftmargin=25pt}
%\setenumerate[0]{leftmargin=25pt}




\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
}
\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={ ()},
             pdfkeywords = {},  
            pdftitle={DRAFT - PS 811: Introduction to Statistical Computing},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls


\setcounter{secnumdepth}{0}





\usepackage{setspace}

\title{DRAFT - PS 811: Introduction to Statistical Computing}
\author{Marcy Shieh}
\date{Fall 2020}


\begin{document}  

		\maketitle
		
	
		\thispagestyle{firststyle}

%	\thispagestyle{empty}


	\noindent \begin{tabular*}{\textwidth}{ @{\extracolsep{\fill}} lr @{\extracolsep{\fill}}}


E-mail: \texttt{\href{mailto:mshieh2@wisc.edu}{\nolinkurl{mshieh2@wisc.edu}}} & Web: \href{http://marcyshieh.github.io}{\tt marcyshieh.github.io}\\
Office Hours: TBD  &  Class Hours: TBD\\
Office: polisci811-fall2020.slack.com  & Class Room: TBD\\
	&  \\
	\hline
	\end{tabular*}
	
\vspace{2mm}
	


\hypertarget{course-description}{%
\section{Course Description}\label{course-description}}

Much of modern social science takes place on a computer. As you mature
in the discipline, you will find yourself desperate for efficiency and
organization. This class will guide both quantatiative and qualitative
scholars on how to better organize data, conduct analyses, compile
literature reviews, create bibliographies, and complete a paper using
popular computational tools and reproducible workflows. While many
research projects are inundated with decisions that require theoretical
justification, this class will at least help you make your analytical
work transparent to the casual interloper.

Political Science 811 is an introductory class of computing resources
for political and social sciences. While not all political scientists
use these tools (yet), most of the top journals in the discipline (e.g.,
\emph{Journal of Politics, American Political Science Review, American
Journal of Political Science}, and other journals that you may want to
publish in) now require authors to submit replication files. Instead of
having to retrace all your steps, these tools will allow you to organize
as you go along and keep you honest as a scholar. In addition, I can
attest that these skills transfer to technical and research-oriented
jobs outside of academia. Even if your job description is not strictly
technical in nature, you may end up having to interact with software
developers and data analysts who operate within these computing
frameworks.

The goal of this course is to familiarize you with popular computing
tools and where the discipline is headed so you can implement them in
your own work. Other method courses focus on the foundational aspects of
quantitative and qualitative research, which are important and great,
but they leave a gap between ``doing the research'' and ``writing the
paper.'' This class is meant to fill in that gap, which can often be
complicated, confusing, and ambiguous. The hope is that this class will
introduce you to tools that will ameliorate that daunting chasm, or at
least make it more managaeable.\footnote{Special thanks to my PS 811
  predecessors, \href{https://mikedecr.github.io/}{Mike DeCrescenzo} and
  \href{https://judgelord.github.io}{Devin Judge-Lord}, for sharing
  their materials with me.}

\hypertarget{learning-outcomes}{%
\section{Learning Outcomes}\label{learning-outcomes}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Prepare, model, and visualize data using the statistical programming
  language \textbf{R}.
\item
  Write a research paper using \textbf{R Markdown}.
\item
  Manage your research projects using \textbf{Git}.
\end{enumerate}

\hypertarget{course-credit-information}{%
\section{Course Credit Information}\label{course-credit-information}}

This class is a 1-credit pass/fail course. This means that the class
meets for one (1) hour each week over the fall semester and carries the
expectation that students will work on course materials for about two
(2) hours out of the classroom for every class period.

Our course will meet for two (2) hours every week. Students
\emph{should} attend the first hour synchronously (TBD) or
asynchronously to gain all the necessary materials for the course. This
hour will include lectures, code demos, and some applied exercises.
While there may be some code snippets on the lecture slides, I will
upload the full code scripts to Canvas so you can reference them later.

Student then have the option to attend the second hour, where I will go
over some questions on the problem set and answer any burning questions
you may have about the course materials. In essence, you may interpret
the second hour as additional office hours.

\hypertarget{required-software}{%
\section{Required Software}\label{required-software}}

You will need to download the following software tools on to your
computer. Fortunately, they are all free.

\begin{itemize}
\item
  \textbf{R}, a free software environment to conduct statistical
  analyses and create data visualizations. For those uninitiated to the
  world of statistical programming, R is basically a super powerful
  version of Microsoft Excel, but way more efficient. Download R at
  \url{https://www.r-project.org/}.
\item
  \textbf{R Studio}, a user-friendly environment for programming in R.
  One way to think about this is that R is the engine, and R Studio is
  the car. You can use the engine without the car, but you cannot drive
  the car without the engine. Or something like that. Downloda RStudio
  Desktop (the free open source version) at
  \url{https://rstudio.com/products/rstudio/download/}.
\item
  \textbf{LaTeX}, a software system for preparing documents. I don't
  envision us using LaTeX in its natural state, but you will need to
  download the software in order to compile R Markdown documents.
  Download LaTeX at \url{https://www.latex-project.org/get/}.
\item
  \textbf{Git}, a version control system. You know about track changes
  in Microsoft Word? Git is track changes, but for all your files. It
  tells you who made changes to the file, what changes they made, and
  when they made them. Thus, it is ideal for collaboration and working
  solo. Instead of naming your files \texttt{document\_2020-08-25.Rmd}
  then over-writing it as \texttt{document\_2020-08-26.Rmd} when you
  make some changes, Git does all of this for you without you possibly
  confusing dates, saving the file without saving \emph{as} the file,
  and all the other problems you may run into when you attempt to
  version control manually. You may already have Git in your computer.
  Check whether you have Git on your computer at
  \url{https://git-scm.com/book/en/v2/Getting-Started-Installing-Git}.
  \textbf{Do not install from source.}
\end{itemize}

I am not dogmatic about how you choose to code in R, as long as you
\emph{know} how to code and \emph{understand} what you are doing. Half
the battle with learning R is fostering confidence and curiosity. Only
then, would you feel comfortable running different commands and Googling
your R questions.

This class is an introduction on how to use R, and aims to give you the
ability to self-learn R and be literate in all kinds of R code. Thus, I
will try to cover both base R and the Tidyverse language, which I will
talk about more during the course, but it is ultimately up to you to
determine what works for you.

\hypertarget{deliverables-grading}{%
\section{Deliverables \& Grading}\label{deliverables-grading}}

The assignments in this class are meant to give you extra practice on
the materials we learn in class and contribute to your final project.
Late assignments will not be accepted.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Weekly assignments: 50\%} These assignments will help you
  apply the skills learned in class and will focus on data tasks in R,
  then integrating them into the Git workflow.
\item
  \textbf{Midterm project: 20\%} You and a partner will work on a
  project together where you demonstrate your ability to use Git for
  collaboration purposes. You will present your project live or submit a
  recording. Your colleagues can provide feedback on Slack.
\item
  \textbf{Final project: 30\%} You will be responsible for submitting a
  project repository that includes the R Markdown paper, the R code for
  the analysis, graphics, and bibliography, etc. and is
  version-controlled with Git and pushed to Github. In essence, I should
  be able to reproduce your entire project with the files in your
  project respository. You will present your project live or submit a
  recording. Your colleagues can provide feedback on Slack.
\end{enumerate}

\hypertarget{class-logistics}{%
\section{Class Logistics}\label{class-logistics}}

\hypertarget{attendance}{%
\subsection{Attendance}\label{attendance}}

Attendance is not required but strongly recommended. If you cannot
attend class synchronously, you should watch the video recordings of the
lectures in your own time. The advantage of attending class
synchronously is that you will be \emph{encouraged} to ask questions as
you learn the concepts. However, if it is not possible at all for you to
attend lecture synchronously (e.g., life circumstances, time zone
differences), please let me know and we can discuss accommodation
options.

\hypertarget{academic-integrity}{%
\subsection{Academic Integrity}\label{academic-integrity}}

Academic integrity is important to the University of Wisconsin--Madison.
As such, please do not plagiarize, use unauthorized materials or
fabricated data, tamper with another student's work, or assist any
students in these acts. If you collaborated with your colleagues (i.e.,
worked on the assignment together), do not copy and paste their code or
pass them as your own. To keep yourself honest, you may also link to the
website(s) or forum post(s) that helped you find the answer. Type out
the code and credit them for their help. Explain the code in ways that
you understand and can reference in the future. If you fail to follow
best practices, you can face disciplinary action, which includes failing
the assignment, failing the course, or facing other disciplinary action
by the department or the university. I will report substantial or
repeated cases of misconduct to the Office of Student Conduct \&
Community Standards for further review. Refer to
\url{https://conduct.students.wisc.edu/misconduct/academic-integrity/}
for more information.

\hypertarget{diversity-inclusion}{%
\subsection{Diversity \& Inclusion}\label{diversity-inclusion}}

Diversity is a source of strength, creativity, and innovation for
UW--Madison. We value the contributions of each person and respect the
profound ways their identity, culture, background, experience, status,
abilities, and opinion enrich the university community. We commit
ourselves to the pursuit of excellence in teaching, research, outreach,
and diversity as inextricably linked goals. The University of
Wisconsin--Madison fulfills its public mission by creating a welcoming
and inclusive community for people from every background --- people who
as students, faculty, and staff serve Wisconsin and the world. Visit
\href{https://diversity.wisc.edu/}{diversity.wisc.edu} for more
information.

\hypertarget{accommodations}{%
\subsection{Accommodations}\label{accommodations}}

The University of Wisconsin-Madison supports the right of all enrolled
students to a full and equal educational opportunity.

If you observe any religious holidays, please let me know and we can
discuss accommodations.

The Americans with Disabilities Act (ADA), Wisconsin State Statute
(36.12), and UW-Madison policy (Faculty Document 1071) require that
students with disabilities be reasonably accommodated in instruction and
campus life. Reasonable accommodations for students with disabilities is
a shared faculty and student responsibility. Students are expected to
inform faculty {[}me{]} of their need for instructional accommodations
by the end of the third week of the semester, or as soon as possible
after a disability has been incurred or recognized. I will work either
directly with the student you or in coordination with the McBurney
Center to identify and provide reasonable instructional accommodations.
Disability information, including instructional accommodations as part
of a student's educational record, is confidential and protected under
FERPA. Visit \url{https://mcburney.wisc.edu/instructor/} for more
information.

Sudden life circumstances can certainly impact class performance. There
are various resources on campus to help you navigate difficult times,
such as the \href{https://doso.students.wisc.edu/}{Dean of Students
Office} and the \href{https://students.wisc.edu/}{Division of Student
Life}.

Please let me know as soon as possible if you require any accommodations
at all, including anything that I did not mention. I want to ensure that
this course contributes to your knowledge base rather than be a
hindrance to your success.

\hypertarget{helpful-resources}{%
\section{Helpful Resources}\label{helpful-resources}}

At some point, you might get stuck, as we all do. But have no fear! Not
only will this class allow you to make mistakes, you will have me and
your fellow colleagues as resources. And, since computing tools are so
ubiquitious in academic and professional circles, it is very likely that
someone also has the same question as you do and asked it on an Internet
forum. With any luck, someone may have even answered the question!

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Class Attendance} You should attend both first and second
  hours of class. The first hour introduces you to concept. The second
  hour allows you to ask questions about what you just learned, making
  it the best way to ask questions and learn.
\item
  \textbf{Class Materials} I strongly recommend that you reference the
  class materials first. Rarely would I include a question in an
  assignment that you will be unable to solve by applying concepts that
  we went over in lecture. The answer keys will reference the slide
  where the concept is discussed in case you want to go back to the
  exact slide and review the concept.
\item
  \textbf{Collaboration} As we continue to navigate our virtual higher
  education landscape, we are confronted with hurdles when it comes to
  ease of collaboration. Even in normal times, however, we often have to
  work with collaborators who we cannot meet in-person. As such, I
  highly encourage you to collaborate with each other on weekly
  assignments. For effective collaboration, I recommend doing the
  exercises on your own, identifying the questions you need help on,
  then showing up to a collaborative space with what you know and what
  you don't know. Collaboration \emph{does} mean seeking help among each
  other, but it \emph{does not} mean copying and pasting text and code
  written by someone else and passing it off as your own. If you use
  your colleague's code, as one often do when they learn coding, type
  out the code yourself and make sure to \emph{credit} your colleague
  for their work. If you have some experience with coding or feel
  comfortable with the material, please consider assisting your
  colleagues. Sharing knowledge is as rewarding as gaining knowledge,
  but hoarding knowledge is not rewarding at all.
\item
  \textbf{Slack} \href{https://slack.com/}{Slack} is a communication
  platform that came to fame among software developers, but is now used
  by project teams of all stripes. Posting on Slack is not a
  requirement, but it is certainly a way for you to reach out to help
  from me and/or your colleagues and get an answer as soon as possible.
  If you believe your question will be beneficial to not only you, but
  to your colleagues as well, you should \emph{definitely} post it on
  the Slack. As a graduate student, I'm in front of my computer quite a
  bit, but I will give you all a chance to help each other with
  questions from the assignment before I jump in. Some examples of
  Slack-friendly items include: questions about the assignments,
  resources that you have found useful on and off the web (e.g., books,
  websites, blogs, Twitter posts, videos), mnemonics, or
  cross-references with your other courses.
\item
  \textbf{Office Hours} While I plan to be available on Slack on most
  weekdays (and maybe even weekends), I will also hold regular office
  hours where I will definitely be available unless otherwise noted. To
  schedule office hours with me, simply send me a direct Slack message
  during my regular office hours and I will call you via Slack once I am
  available.
\item
  \textbf{Email} If your question is of a personal nature and is
  specific to you and only you, please email me. Expect me to respond
  within 24 hours and if I do not, please feel free to email me again.
  You will \emph{not} be annoying me.
\item
  \textbf{Google (or your preferred search engine)} There are
  \emph{tons} of resources online about R and Git. You will start to
  notice that your search results will often yield either posts from
  Stack Overflow and Stack Exchange. Use them. Much of Goggling in R to
  achieve required results is learning how to refer to certain commands.
  You learn as you go. Sometimes the Stack Overflow and Stack Exchange
  communities can get a bit abrasive (in my experience). I recommend the
  \href{https://community.rstudio.com/}{R Studio Community} if you do
  actually want to ask a question because you can't find an answer to it
  anywhere. I love Googling and I have spent many hours going down on
  various R-related Googling rabbit holes late into the night. I don't
  recommend this. While it has helped me learn \emph{how} to Google,
  there comes a point (it's the 5-hour mark for me, but it might be the
  2-hour mark for you) where you should ask for help.
\end{enumerate}

\hypertarget{class-schedule}{%
\section{Class Schedule}\label{class-schedule}}

The schedule includes topics, resources, and due dates for the midterm
and final projects. I will try to keep resources at a minimum. You will
likely be overwhelmed by the R you will learn in PS 812 and the R you
will learn in here, so I want to minimize the confusion and maximize the
practicality.

I will try to provide some instructions for setting up everything you
need prior to each lecture, but feel free to ask questions in class, on
the Slack, or via email. I would rather give you time to set up
everything in your own time and ask for help when convenient than feel
left behind in class and freak out, which is often the story of my life.

This syllabus might evolve over time and accommodate any feedback I
receive over the course of the semester. Think of this syllabus as a
road map. There are certain destinations we'd like to hit, but it's
possible for us to take a different route to get there.

\hypertarget{week-01-0815---0819-set-up-all-the-tools-you-will-need-for-the-class-including-r-r-studio-r-markdown-knit-as-pdf-and-html-git-and-github.}{%
\subsection{Week 01, 08/15 - 08/19: Set up all the tools you will need
for the class, including R, R Studio, R Markdown (knit as PDF and HTML),
Git, and
Github.}\label{week-01-0815---0819-set-up-all-the-tools-you-will-need-for-the-class-including-r-r-studio-r-markdown-knit-as-pdf-and-html-git-and-github.}}

\begin{itemize}
\tightlist
\item
  \href{https://happygitwithr.com/rstudio-git-github.html}{Connect
  RStudio to Git and Github}
\item
  \href{https://bookdown.org/yihui/rmarkdown/markdown-syntax.html}{Markdown
  Syntax}
\end{itemize}

\hypertarget{week-02-0822---0826-do-you-what-you-already-do-but-with-cool-computing-tools-create-a-project-create-a-document-write-some-math-in-your-document-and-organize-a-bibliography.}{%
\subsection{Week 02, 08/22 - 08/26: Do you what you already do but with
cool computing tools: create a project, create a document, write some
math in your document, and organize a
bibliography.}\label{week-02-0822---0826-do-you-what-you-already-do-but-with-cool-computing-tools-create-a-project-create-a-document-write-some-math-in-your-document-and-organize-a-bibliography.}}

\begin{itemize}
\tightlist
\item
  \href{https://swcarpentry.github.io/r-novice-gapminder/02-project-intro/}{Project
  Management with R Studio}
\item
  \href{https://rpruim.github.io/s341/S19/from-class/MathinRmd.html\#:~:text=Math\%20inside\%20RMarkdown,10n\%3D1n2.}{Mathematics
  in R Markdown}
\item
  \href{https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html}{Bibliographies
  and citations}
\end{itemize}

\hypertarget{week-03-0829---0902-base-r-vs.-tidyverse-and-how-to-be-comfortable-with-both.}{%
\subsection{Week 03, 08/29 - 09/02: Base R vs.~Tidyverse, and how to be
comfortable with
both.}\label{week-03-0829---0902-base-r-vs.-tidyverse-and-how-to-be-comfortable-with-both.}}

\begin{itemize}
\tightlist
\item
  \href{https://tavareshugo.github.io/data_carpentry_extras/base-r_tidyverse_equivalents/base-r_tidyverse_equivalents.html}{Syntax
  equivalents: base R vs Tidyverse}
\end{itemize}

\hypertarget{week-04-0905---0909-some-useful-functions-for-your-own-work-and-ps-812.}{%
\subsection{Week 04, 09/05 - 09/09: Some useful functions for your own
work and PS
812.}\label{week-04-0905---0909-some-useful-functions-for-your-own-work-and-ps-812.}}

\begin{itemize}
\tightlist
\item
  \href{https://r4ds.had.co.nz/iteration.html}{Iterations} and
  \href{https://r4ds.had.co.nz/many-models.html}{Many Models}
\item
  \href{https://www.hvitfeldt.me/blog/purrr-tips-and-tricks/}{Purrr -
  tips and tricks}
\end{itemize}

\hypertarget{week-05-0912---0916-manipulate-data.}{%
\subsection{Week 05, 09/12 - 09/16: Manipulate
data.}\label{week-05-0912---0916-manipulate-data.}}

\begin{itemize}
\tightlist
\item
  \href{https://mikedecr.github.io/courses/811/811-03-data/}{Data
  Manipulation}
\end{itemize}

\hypertarget{week-06-0919---0923-create-graphics.}{%
\subsection{Week 06, 09/19 - 09/23: Create
graphics.}\label{week-06-0919---0923-create-graphics.}}

\begin{itemize}
\tightlist
\item
  \href{https://mikedecr.github.io/courses/811/811-04-graphics/}{Graphics}
\end{itemize}

\hypertarget{week-07-0926---0930-regression-and-model-outputs.}{%
\subsection{Week 07, 09/26 - 09/30: Regression and model
outputs.}\label{week-07-0926---0930-regression-and-model-outputs.}}

\begin{itemize}
\tightlist
\item
  \href{https://mikedecr.github.io/courses/811/811-05-analysis/}{Analysis}
\item
  \href{https://raw.githack.com/uo-ec607/lectures/master/08-regression/08-regression.html}{Regression
  Analysis in R}
\end{itemize}

\hypertarget{week-08-1003---1007-midterm-project-presentations.}{%
\subsection{Week 08, 10/03 - 10/07: Midterm project
presentations.}\label{week-08-1003---1007-midterm-project-presentations.}}

\hypertarget{week-09-1010---1014-dealing-with-data.}{%
\subsection{Week 09, 10/10 - 10/14: Dealing with
data.}\label{week-09-1010---1014-dealing-with-data.}}

\hypertarget{week-10-1017---1021-popular-analysis-tools-text-analysis-mturkr.}{%
\subsection{Week 10, 10/17 - 10/21: Popular analysis tools: text
analysis,
mturkr.}\label{week-10-1017---1021-popular-analysis-tools-text-analysis-mturkr.}}

\hypertarget{week-11-1024---1028-some-more-git.}{%
\subsection{Week 11, 10/24 - 10/28: Some more
git.}\label{week-11-1024---1028-some-more-git.}}

\hypertarget{week-12-1031---1104-create-your-own-github-website.}{%
\subsection{Week 12, 10/31 - 11/04: Create your own Github
website.}\label{week-12-1031---1104-create-your-own-github-website.}}

\begin{itemize}
\tightlist
\item
  \href{https://marcyshieh.github.io/posts/github-website-tutorial/}{How
  to Host a Website on Github}
\end{itemize}

\hypertarget{week-13-1107---1111-final-project-presentations.}{%
\subsection{Week 13, 11/07 - 11/11: Final project
presentations.}\label{week-13-1107---1111-final-project-presentations.}}




\end{document}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother
